Implications of afforestation policies for global markets

### 2023-05-31 paper.Rmd
Update paper.Rmd

# Introduction
Forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation. In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019]. This is because of forests proven ability to sequester carbon not only the stems of trees [@davis2022] but their soils as well [@guo2021].  At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland. Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017]. Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community.

@@

South America and Sub-Saharan Africa are two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70 percent of the committments coming from Sub-Saharan nations [@auda2023]. However, while many of these nations may be well-suited in big-physical terms for afforestation, they also may be politically unstable or border regions of instability. This creates concern about how long-lasting any investment may be. Other nations that have a big potential for gains from afforestation face different challenges. India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers wastelands that may not be well suited for afforestation [@amjath2019].
These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].

There is growing attention being placed globally to reforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].

>>>>

# Introduction
In 1995, there was 4.2 billion hectares of forestland globally,which covered roughly 32.8% of the land [@Humpenöder2014]. These forests play an important role in maintaining and growing the land-based carbon sink, currently sequestering about 30% of GHG emmissions [@roe2019]. With increasingly negative climate impacts, the pressure to enhance the terrestrial sink is growing, and one of the tools to achieve this is through afforestation. In fact, many of the Nationally Determined Contributions (NDCs)  countries have submitted that detail their plans to keep warming under the 2C threshold rely to a significant degree on afforestation projects [@roe2019]. This is because of forests proven ability to sequester carbon not only the stems of trees [@davis2022] but their soils as well [@guo2021].  At the same time, forests are continually threatened by issues such as rising incomes and food prices and population growth that put pressure on policymakers to allow greater conversion of forestland into farmland. Previous research though has suggested that substantive gains in the carbon sink (+23.8PgCo2e/yr) can be made without sacrificing agricultural land [@griscom2017]. Thus, the manner in which these various and oftentimes competing concerns are addressed will have significant impacts on the global community. There is growing attention being placed globally to afforestation efforts, which is highligted by the UN's declaration of the years from 2021 through 2030 as the UN Decade of Ecosystem Resotration [@sewell2020].

@@

South America and Sub-Saharan Africa are two regions that show much potential for afforestation. It is estimated that these two regions hold at least 50% of the potential global gain while regions such as Northern Africa and the Middle East show little promise as growth rates are quite low [@doelman2020]. Thus, it is promising that the African Forest Landscape Restoration Initiative (AFR100) aims to afforest 100 million hectares by 2030 with roughly 70 percent of the committments coming from Sub-Saharan nations [@auda2023]. While many of these nations may be well-suited in big-physical terms for afforestation, there is concern about how long-lasting any investment may be due to proximate zones of political instability. Other nations that have a big potential for gains from afforestation face different challenges. India has a great potential to increase its carbon sequestration through afforestation, but due to its masssive population and food security worries, its NDC does not entertain the idea of agricultural land being repurposed and instead primarily considers targeting wastelands that may not be well suited for afforestation [@amjath2019].

These are not the only concerns that need to be considered by policymakers planning afforestation activities, as for example, oelman2020].
